\textbf{Are regime changes always bad economics? Evidence from daily financial data}

### 2023-07-04 doc/regime_changes_bp_rr.tex
Updates to rr manuscript

We also provide methodological advancements when compared to the aforementioned cross country studies. Specifically, we conduct event studies of daily financial data, which estimate a local average treatment effect of an unexpected event on stock prices \textit{at the exact time of the event}. This interrupted time-series approach mitigates the endogeneity problems in previous cross country regressions---confounding events would need to occur on the same day as instability, and do so for a large portion of all of our independently tested events in order to influence our estimates.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.][]{fisman2001estimating,faccio2006politically,goldman2009politically}. Studies in ``forensic economics" have used abnormal returns from political events to locate or examine transactions such as insider trading \citep{dube2011coups}, illegal arms trading \citep{dellavigna2010detecting}, and the impact of hostilities on the financial performance of diamond mining firms in Angola \citep{guidolin2007diamonds}.} In addition, we integrate synthetic control and event study methods and demonstrate how ``synthetic'' control portfolios can be used to apply market-model event studies even when a suitable control portfolio is absent.

@@

\textcolor{red}{The key identifying assumption behind our interrupted time-series approach is that the events are unpredicted in an immediate temporal sense. In other words, for an event to be plausibly exogenous, the exact timing of the event must be unknown to investors. If information about the event has leaked prior to occurrence of the event, a portion of the change in investment will already be priced into the stock during the estimation window.}

@@

A constant mean return model is used instead of a market model in order to maximize the number of observations and because our unit of analysis is country-wide market indices rather than firms.\footnote{Plausible market indices such as the MSCI World Index and the S\&P/IFC Emerging Markets Investable Composite Index only begin in 1976 and 1995, respectively.} To address concerns regarding use of a constant mean return model, we also create a synthetic control portfolio for each event and compare observed returns to the synthetic returns.

@@

\autoref{fig:volatility} shows the mean volatility ($\overline{\sigma_t}$) estimates from the GARCH (1,1) model across all irregular regime changes for 250 trading days prior to and 250 days after each event. As expected, the volatility estimates stay between a narrow range at nearly all dates except those surrounding the regime change. Volatility appears to increase slowly just before the regime change, albeit not to a degree out of line with previous fluctuations in volatility. This may suggest that investors sometimes have information about the events before they occur. Nonetheless, there is still an enormous volatility jump on the day of the regime change. Volatility then decreases to normal levels within a month of the event.

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those in \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven by five events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; the assassination of Indian Prime Minister Indira Gandhi on October 31, 1984; the suicide bombing that killed Sri Lankan president Ranasinghe Premadasa on May 1, 1993; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

@@

Second, we create a synthetic control portfolio for each event based on the techniques introduced in \citet{abadie2003economic} and \citet{abadie2010synthetic}. \textcolor{red}{This approach creates a ``synthetic'' time series of daily financial returns based on the returns in a basket of countries which did \textit{not} experience tumultuous political events on the day of the event. We are then able to compare the returns in the country which experienced the unexpected regime change with the actual returns of a reference market---in this case all countries that did not experience an unexpected regime change.} Each country is given a weight which represents its influence in the synthetic control portfolio. The weight is chosen so that the daily returns and the variance of the daily returns of the control portfolio and the event country are most similar in the estimation window.\footnote{See ``Synthetic Control Portfolio'' in the Appendix for mathematical formalization.} The set of possible countries in the control portfolio consists of all countries listed in \autoref{tab:stock-list}. \textcolor{red}{In contrast with the large and significant abnormal returns in the actual countries of study, the returns in the synthetic control portfolio are [add percentage difference in magnitude] and never significantly different from zero for coups, assassinations, or resignations (or overall).}

@@

\textcolor{red}{Our findings have a number of potential implications for policymakers. First, our finding that even democratically subversive irregular regime changes can have a positive impact on investment when the replacement government is clearly pro-business  implies that investor goals and actions may be at odds with regime stability. Second, the immediate nature and large magnitude of the financial outflows we observe in the wake of the majority of coups and assassinations suggests that a weakened financial system and economy are likely to accompany these kinds of regime changes. More optimistically, however, we also find suggestive evidence that in the event of a democratic extra-legal regime change, clearly signalling a commitment to democracy may result in increased investment in a country's firms.}

There are a number potential avenues for future research. First, more research is needed to identify the pro and anti market characteristics of regime changes. There may be pro or anti market features of regime changes that are more nuanced than those identified in this paper. Similarly, it would be useful to identify additional mechanisms through which regime changes effect investment. In the case of Chile,  \citet{girardi2018institution} show that increasing stock market valuations were caused by changes in private property rights rather than economic growth prospects or wage costs. Research focusing on the mechanisms in other settings would help generalize these findings. Lastly, it would be helpful to determine the extent to which stock market returns translate to broader economic development outcomes. \citet{meyersson2016political} has made progress on this front by examining the impact of coups on a number of outcomes in addition to economic growth including investment, debt, inflation, infant mortality, and years of schooling. It would be fruitful to examine whether the direction of the effects of different types of regime changes on these outcomes are consistent with their stock market effects, or if investor perceptions are at odds with certain development goals. Such research can help enlarge the body of evidence on the extent to which regime changes cause institutional and political change, and, in turn, have significant consequences for economic development.

@@

\begin{figure}[!ht]
\includegraphics[scale=0.8]{../figs/bangla-1990.pdf}
\caption{CARs during the 1998 Indonesian revolution}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\includegraphics[scale=0.8]{../figs/indonesia-1998.pdf}
\caption{CARs during the 1998 Indonesian revolution}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\includegraphics[scale=0.8]{../figs/phil-2001.pdf}
\caption{CARs during the 1998 Indonesian revolution}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\includegraphics[scale=0.8]{../figs/argentina-2001.pdf}
\caption{CARs during the 2001 Argentinian riots}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\includegraphics[scale=0.8]{../figs/ukraine-2004.pdf}
\caption{CARs during the 2004 Orange Revolution}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\includegraphics[scale=0.8]{../figs/ecuador-2005.pdf}
\caption{CARs during the 2006 Nepalese Democracy Movement}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\includegraphics[scale=0.8]{../figs/nepal-2006.pdf}
\caption{CARs during the 2006 Nepalese Democracy Movement}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\includegraphics[scale=0.8]{../figs/tunisia-2011.pdf}
\caption{CARs during the 2011 Tunisian Revolution}
\label{fig:CAR-Egypt}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[scale = 0.75]{../figs/mean-volatility-egarch.pdf}
\caption{Mean of volatility estimates from exponential GARCH models}
\label{fig:volatility}
\end{figure}

\begin{figure}[H]
\includegraphics[scale = 0.75]{../figs/mean-volatility-tgarch.pdf}
\caption{Mean of volatility estimates from threshold GARCH models}
\label{fig:volatility}
\end{figure}

>>>>

We also provide methodological advancements when compared to the aforementioned cross country studies. Specifically, we conduct event studies of daily financial data, which estimate a local average treatment effect of an unexpected event on stock prices \textit{at the exact time of the event}. This interrupted time-series approach mitigates the endogeneity problems in previous cross country regressions---confounding events would need to occur on the same day as instability, and do so for a large portion of all of our independently tested events in order to influence our estimates.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.][]{fisman2001estimating,faccio2006politically,goldman2009politically}. Studies in ``forensic economics" have used abnormal returns from political events to locate or examine transactions such as insider trading \citep{dube2011coups}, illegal arms trading \citep{dellavigna2010detecting}, and the impact of hostilities on the financial performance of diamond mining firms in Angola \citep{guidolin2007diamonds}.} In addition, we integrate synthetic control and event study methods and demonstrate how ``synthetic'' control portfolios can be used to apply a method akin to market-model event studies even when a suitable control portfolio is absent.

@@

\textcolor{red}{The key identifying assumption behind our interrupted time-series approach is that the events are unpredicted in an immediate temporal sense. In other words, for an event to be plausibly exogenous, the exact timing of the event must be unknown to investors. If information about the event has leaked prior to occurrence of the event, a portion of the change in investment will already be priced into the index during the estimation window.}

@@

A constant mean return model is used instead of a market model in order to maximize the number of observations and because our unit of analysis is country-wide market indices rather than firms.\footnote{Plausible market indices such as the MSCI World Index and the S\&P/IFC Emerging Markets Investable Composite Index only begin in 1976 and 1995, respectively.} To address concerns regarding use of a constant mean return model, we also create a synthetic control portfolio for each event and compare observed returns to the synthetic returns as a robustness check.

@@

\autoref{fig:volatility} shows the mean volatility ($\overline{\sigma_t}$) estimates from the GARCH (1,1) model across all irregular regime changes for 250 trading days prior to and 250 days after each event.\footnote{Robustness checks using exponential GARCH and threshold GARCH models can be found in \autoref{fig:egarch} and \autoref{fig:tgarch}, respectively.} As expected, the volatility estimates stay between a narrow range at nearly all dates except those surrounding the regime change. Volatility appears to increase slowly just before the regime change, albeit not to a degree out of line with previous fluctuations in volatility. This may suggest that investors sometimes have information about the events before they occur. Nonetheless, there is still an enormous volatility jump on the day of the regime change. Volatility then decreases to normal levels within a month of the event.

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those in \autoref{tab:AR-coups} but for assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven by five events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; the assassination of Indian Prime Minister Indira Gandhi on October 31, 1984; the suicide bombing that killed Sri Lankan president Ranasinghe Premadasa on May 1, 1993; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

@@

Second, we create a synthetic control portfolio for each event based on the techniques introduced in \citet{abadie2003economic} and \citet{abadie2010synthetic}. \textcolor{red}{This approach creates a ``synthetic'' time series of daily financial returns based on the returns in a basket of countries which did \textit{not} experience tumultuous political events on the day of the event. We are then able to compare the returns in the country which experienced the unexpected regime change with the actual returns of a reference market---in this case all countries that did not experience an unexpected regime change.} Each country is given a weight which represents its influence in the synthetic control portfolio. The weight is chosen so that the daily returns and the variance of the daily returns of the control portfolio and the event country are most similar in the estimation window.\footnote{See ``Synthetic Control Portfolio'' in the Appendix for mathematical formalization.} The set of possible countries in the control portfolio consists of all countries listed in \autoref{tab:stock-list}. \textcolor{red}{In contrast with the large and significant abnormal returns in the actual countries of study, the returns in the synthetic control portfolio are never significantly different from zero for coups, assassinations, or resignations (or overall).}

@@

\textcolor{red}{Our findings have a number of potential implications for policymakers. First, our finding that even democratically subversive irregular regime changes can have a positive impact on investment when the replacement government is clearly pro-business  implies that investor goals and actions may be at odds with regime stability. Second, the immediate nature and large magnitude of the financial outflows we observe in the wake of the majority of coups and assassinations suggests that a weakened financial system and economy are likely to accompany these kinds of regime changes. More optimistically, however, we also find suggestive evidence that in the event of a democratic regime change or revolution, clearly signalling a commitment to democracy may result in increased investment in a country's firms.}

There are a number potential avenues for future research. First, more research is needed to identify the pro and anti market characteristics of regime changes. There may be pro or anti market features of regime changes that are more nuanced than those identified in this paper. Similarly, it would be useful to identify additional mechanisms through which regime changes effect investment. In the case of Chile,  \citet{girardi2018institution} show that increasing stock market valuations were caused by changes in private property rights rather than economic growth prospects or wage costs. Research focusing on the mechanisms in other settings would help generalize these findings. \textcolor{red}{Second, it remains unclear how markets react to successful irregular regime change conditional on attempt. A future study could therefore examine how markets react to successful assassinations or coups when compared to failed attempts.} Lastly, it would be helpful to determine the extent to which stock market returns translate to broader economic development outcomes. \citet{meyersson2016political} has made progress on this front by examining the impact of coups on a number of outcomes in addition to economic growth including investment, debt, inflation, infant mortality, and years of schooling. It would be fruitful to examine whether the direction of the effects of different types of regime changes on these outcomes are consistent with their stock market effects, or if investor perceptions are at odds with certain development goals. Such research can help enlarge the body of evidence on the extent to which regime changes cause institutional and political change, and, in turn, have significant consequences for economic development.

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.8]{../figs/bangla-1990.pdf}
\caption{CARs during the 1990 Bangladesh uprising}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.8]{../figs/indonesia-1998.pdf}
\caption{CARs during the 1998 Indonesian revolution}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.8]{../figs/phil-2001.pdf}
\caption{CARs during the Philippines Second People Power Revolution}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.8]{../figs/argentina-2001.pdf}
\caption{CARs during the 2001 Argentine crisis}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.8]{../figs/ukraine-2004.pdf}
\caption{CARs during the 2004 Orange Revolution}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.8]{../figs/ecuador-2005.pdf}
\caption{CARs during the Ecuadorian Revolution of 2005}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.8]{../figs/nepal-2006.pdf}
\caption{CARs during the 2006 Nepalese Democracy Movement}
\label{fig:CAR-Egypt}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.8]{../figs/tunisia-2011.pdf}
\caption{CARs during the 2011 Tunisian Revolution}
\label{fig:CAR-Egypt}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[scale = 0.75]{../figs/mean-volatility-egarch.pdf}
\caption{Mean of volatility estimates from exponential GARCH models}
\label{fig:egarch}
\end{figure}

\begin{figure}[H]
\includegraphics[scale = 0.75]{../figs/mean-volatility-tgarch.pdf}
\caption{Mean of volatility estimates from threshold GARCH models}
\label{fig:tgarch}
\end{figure}
